### 2018-03-30 reports/phylo_report.rmd
not much to report at the moment

---
title: "phylo_analyses"
author: "Everyone"
date: "3/22/2018"
output: pdf_document
---

run phylo model

```{r}

mod2<-run_phylo_lm(traits,
                         response_variables,
                         aus_bird_tree)

summary(mod2)
```
maybe we eventually want to do model selection

```{r}
# library(MuMIn)
# dredge(mod2)
```

this is sort of cool.  phylogenetic signal in all the traits

```{r}
  # library(phylosignal)
  # library(phylobase)
  # dd$body_size<-log10(dd$body_size)
  # dd2<-subset(dd,select=-gregariousness)
  # out<-phylo4d(tree_plotting_2,tip.data=dd2)
  # ps<-phyloSignal(out)
  # knitr::kable(ps$stat)
```