### 2018-04-06 reports/phylo_report.rmd
model selection is a nightmare

---
title: "phylo_analyses"
author: "Everyone"
date: "3/22/2018"
output: pdf_document
---

run phylo model

```{r}
traits %>%
    rownames_to_column('names') %>%
    filter(complete.cases(.)) %>%
    column_to_rownames('names') -> trait

  # subset the tree
  matched_tree <- drop.tip(aus_bird_tree,aus_bird_tree$tip.label[!aus_bird_tree$tip.label%in%row.names(trait)])

  response_variables$binom<-response_variables$SCIENTIFIC_NAME_tree
  inner_join(response_variables,trait) %>%
    filter(binom!="not_treated_as_species") %>%
    data.frame(.)-> trait

  dd<-trait
  dd$phylo<-dd$binom
  phylo<-matched_tree # this is hilarious: tree has to be called "phylo" for lmer to work
```

```{r}

dd$lbody<-log10(dd$mean_body_size)
dd$lbrain<-log10(dd$brain_size)

 basic_gaus_fit <- lmer(urban_median~lbody+clutch_size+iucn_status+lbrain+feeding_habitat_generalism+
     breeding_habitat_generalism+granivore+insectivore+carrion_eater+diet_generalism+
     movement_class+ground_nesting+hollow_nesting+nest_generalism+breeding+nest_aggregation+feeding_aggregation+(1|phylo),
                         data=dd,
                         control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),na.action = "na.fail")

 AIC(basic_gaus_fit)
```

Ben Bolker hates model selection (and signficance testing) so no easy way to do it on lmer objects.

for now just test the model that the non-phylo found

```{r}
dd$urban_median_rescale<-(dd$urban_median-mean(dd$urban_median))/sd(dd$urban_median)
dd$clutch_size_rescale<-(dd$clutch_size-mean(dd$clutch_size))/sd(dd$clutch_size)
dd$lclutch<-log(dd$clutch_size)

 basic_gaus_fit_2 <- lmer(urban_median~lclutch+lbody+
                          (1|phylo),
                         data=dd,
                         control=lmerControl(check.nobs.vs.nlev="ignore",check.nobs.vs.nRE="ignore"),
                        na.action = "na.fail")
AIC(basic_gaus_fit_2)
summary(basic_gaus_fit_2)
```

trying with some other software--i think lmer is bugged for phylogenies.

```{r}
 basic_gaus_fit_phylolm <- phylolm(urban_median~lclutch+lbody,phy=phylo,
                         data=dd)

 summary(basic_gaus_fit_phylolm)
 AIC(basic_gaus_fit_phylolm)
```

trying to find another factor that improves the model but it seems tough
```{r}
  basic_gaus_fit_phylolm <- phylolm(urban_median~lclutch+
                                    lbody+feeding_habitat_generalism+
                                    lbrain+insectivore+diet_generalism+
                                    hollow_nesting,
                                    phy=phylo,data=dd)

logLik(basic_gaus_fit_phylolm)
```

lmer is totally bugged for phylo applications.  continue with phylolm next time.

```{r}
 basic_gaus_fit_phylolm_full <- phylolm(urban_median~iucn_status+lbrain+feeding_habitat_generalism+
     carrion_eater+lclutch+lbody+
     movement_class+ground_nesting+nest_generalism+nest_aggregation,
                                    phy=phylo,data=dd)

dredge(basic_gaus_fit_phylolm_full)

```

```{r}

visreg(basic_gaus_fit_phylolm_full)
```