### 2018-04-24 reports/phylo_report.rmd
phylo signal stuff

---
title: "phylo_signal_analyses"
author: "Everyone"
date: "3/22/2018"
output: pdf_document
---

```{r}
require(phylosignal)
   dplyr::select(analysis_data,body_size_logged,clutch_size_logged,response,range_size, #select appropriate rows
          nest_generalism,diet_generalism,breeding_habitat_generalism,feeding_habitat_generalism,brain_residual) ->dd

row.names(dd)<-analysis_data$binom #name rows so that it matches the tree
dd$rand<-rnorm(dim(dd)[2]) #random numbers to test package
p4d <- phylo4d(aus_bird_tree, dd) #create phylobase object

ps<-phyloSignal(p4d,reps = 9999) #run calculation, p values a bit unstable at 999 reps
```
There are two output table the summary statistics and the p-value first the summary stats
```{r}
knitr::kable(ps$stat)
```
and then the p-values
```{r}
knitr::kable(ps$pvalue)
```