### 2018-05-14 scripts/Appendix_5-model_output_results.Rmd
R markdown doc to make Appendix 5 - raw model outputs

---
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
```

## Model: Non-phylogenetically controlled global model results

```{r Non-phylogenetically controlled global model results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary(global_model)
```

## Model: Non-phylogenetically controlled model-averaged results

```{r Non-phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary <- readRDS("Data/dredged_model_summary_results.rds")
summary(summary)
```

## Model: Phylogenetically controlled global model results

```{r Phylogenetically controlled global model results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary(phy_mod_rescaled)
```

## Model: Phylogenetically controlled model-averaged results

```{r Phylogenetically controlled model-averaged results, echo=FALSE, message=FALSE}
setwd("H:/Dissertation/Dissertation Chapters/Data Chapters/Adaptations to urban living in birds/bird_urbanness")
remake::dump_environment()
summary2 <- readRDS("Data/PHYLO_dredged_model_summary_results.rds")
summary(summary2)
```